---
title: "01 Lung Data Report"
author: "Heidi Steiner"
date: "2/21/2022"
output:
  html_document:
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(tidyverse)
library(survival)

ggplot2::theme_set(ggplot2::theme_minimal())
```

## Lung Data 



The lung data set is found in the [`survival`](https://github.com/therneau/survival) R package. Try to load the lung data set in R by issuing the following command at the console `lung`. See the **data dictionary** below: 


|col_name|key|
|:-------|:--|
|inst|Institution code|
|status|	censoring status 1=censored, 2=dead|
|age|	Age in years|
|sex|	Male=1 Female=2|
|time|	Survival time in days|
|ph.ecog|	ECOG performance score (0=good 5=dead)|
|ph.karno|	Karnofsky performance score (bad=0-good=100) rated by physician|
|pat.karno|	Karnofsky performance score as rated by patient|
|meal.cal|	Calories consumed at meals|
|wt.loss|	Weight loss in last six months|

```{r echo = FALSE, message=FALSE}
mockdata = read_csv("2022-02-22_lungdat.csv") %>%  ## read in your data with read_csv
  mutate_at(.vars = c("inst", "status", "sex", "ph.ecog"),.funs = as.factor) 
```

## Easy Exploration of Data 

Many packages, such as [`dplyr`](https://dplyr.tidyverse.org/) facilitate the ease of exploring your data.

```{r comment = ""}
mockdata %>% 
  glimpse()
```

`glimpse()` shows us the dimensions of our dataframe and then the first 19 observations of each column (and the column type (i.e. 'fct' for factor)). 


## Communicate Results Reproducibly

RMarkdown allows you to visualize your data within a document that also contains text. 

```{r warning = F}
mockdata %>% 
  ggplot(aes(time, wt.loss, color = inst)) + 
  geom_point() +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
    )
```



## Summary

- RMD can be used to create data driven reports for your collaborators or even to write entire manuscripts! 
- RMD anatomy includes 
  1. YAML metadata  
  1. markdown text  
  1. code   
  1. output
- RMD is wizardry! 


