---
title: "02 Lung Data Report"
author: "Heidi Steiner"
date: "2/22/2022"
output:
  html_document: 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(cowplot)
library(janitor) # need to install 
library(kableExtra)
library(knitr)
library(skimr)
library(survival)
library(survminer) # need to install
library(tidyverse)

ggplot2::theme_set(ggplot2::theme_minimal())

```


```{r echo = FALSE, message=FALSE}
mockdata = read_csv("2022-02-22_lungdat.csv") %>% 
  mutate_at(.vars = c("inst", "status", "sex", "ph.ecog", "ph.karno", "pat.karno"),.funs = as.factor)
```

## Data Summaries 

Easily print summary tables of your data with [`skimr`](https://docs.ropensci.org/skimr/)

```{r}
mockdata %>%  
  skim()
```


but this is not super reader friendly. ...we can improve table output with `kable()` from the [`knitr`](https://yihui.org/knitr/) package and `kable_styling()` from the [`kableExtra`](https://haozhu233.github.io/kableExtra/) package. 

I used the option `font_size = 9` to reduce the size of the text as well 


```{r}
mockdata %>% 
  skim() %>% 
  knitr::kable() %>% 
  kable_styling(font_size = 9)
```


If you're like me and the NAs in this output are a deal breaker, try splitting the factor variables from the numeric variables so there are fewer columns to report! 

The `.$factor` function will filter just the factor variables from the skim list.

```{r warning=FALSE}
mockdata  %>% 
  skim_to_list() %>% 
  .$factor %>% 
  knitr::kable( caption = 'Factor Variables') %>% 
  kable_styling( font_size = 9)
```

and `.$numeric` function will filter just the numeric variables from the skim list.


```{r warning=FALSE}
mockdata  %>% 
  skim_to_list() %>% 
  .$numeric %>% 
  knitr::kable( caption = 'Numeric Variables') %>% 
  kable_styling( font_size = 9)
```

## Tables 

We also often want to create basic (or not so basic) tables of results for our reports. Create quick frequency tables with [`janitor`](https://garthtarr.github.io/meatR/janitor.html)


```{r comment=""}
mockdata %>% 
  tabyl(ph.ecog)
```

`janitor` also has a mechanism to add frequency percentages and row totals to your tables 

```{r comment=""}
mockdata %>% 
  tabyl(ph.ecog) %>% 
  adorn_totals("row") %>%
  adorn_pct_formatting()
```

and then you can add `kable()` and `kable_styling()` again to make it print how you need 

```{r results='asis'}
mockdata %>% 
  tabyl(ph.ecog, sex) %>%
  knitr::kable(caption = 'ECOG performance score (0=good 5=dead) by Sex (Male=1 Female=2)',
               table.attr = "style='width:30%;'") %>% 
  kable_styling()
```

## Figures 

Often, you want to include results in the form of figures in your reports. [`ggplot2`](https://ggplot2.tidyverse.org/) makes plotting formulaic. 

**Hint** I removed observations with ECOG scores of 3 based on the last table. 

```{r message=FALSE}
p1 = mockdata %>% 
  filter(ph.ecog != "3") %>% 
  ggplot(aes(x = age, y = time, fill = ph.ecog, color = ph.ecog)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(title = "Survival Time by Age of patient within \nECOG performance score groups",
       x = "Age (yrs)",
       y = "Survival Time (days)",
       fill = "ECOG Score \n(0=Good, 2=Less Good)",
       color = "ECOG Score \n(0=Good, 2=Less Good)")
p1 
```

`ggplot2` is super dynamic and includes many ways to visualize your data

```{r}
p2 = mockdata %>% 
  filter(ph.ecog != "3",
         !is.na(ph.karno)) %>% 
  ggplot(aes(x = ph.karno, y = time)) +
  geom_violin(aes(fill = ph.karno,
                  color = ph.karno),
              alpha = .2) +
  geom_boxplot(aes(color = ph.karno),
               outlier.shape = NA,
               width = .1,
               alpha = .5,
               show.legend = F) +
  labs(fill = "0 = Bad\n100 = Good",
       color = "0 = Bad\n100 = Good",
       title = "Karnofsky Performance Score \nrated by Physician",
       x = "Performance Score",
       y = "Time (days)"
       )

p2
```

Or, perhaps you need to include a more domain specific plot. Google for your analysis and R (or tidyverse) and see if this already exists! 

Here, I demonstrate visualizing a survival analysis with the `survival` and [`survminer`](https://rpkgs.datanovia.com/survminer/index.html) packages. 

```{r survminer, fig.height=6}
p3 = ggsurvplot(
    fit = survfit(Surv(time, status) ~ ph.ecog, data = lung %>% filter(ph.ecog != "3")), 
    xlab = "Days", 
    ylab = "Overall survival probability",
    legend.lab = c("0 - Good", "1 - ", "2 - Less Good"),
    legend.title = "ECOG performance score",
    risk.table = T)
p3
```
Sometimes, you'll want to combine multiple plots of different types such as the ones we've just created. [`cowplot`](https://cran.r-project.org/web/packages/cowplot/vignettes/introduction.html) (and `patchwork`) lets us combine multiple plots in a grid. 

First, I align the top plots and then added the survival curve to the grid! 

```{r fig.height=6, fig.width=8, message=FALSE}
cow1 = plot_grid(p1, p2,
          labels = c('A', 'B'), 
          label_size = 12,
          ncol = 2)

plot_grid(cow1, p3$plot, 
          label_size = 12,
          labels = c('', "C"),
          ncol = 1)
```


## Your Turn

#### 1. 

Make this table print nicely when knitted

```{r results='asis'}
mockdata %>% 
  tabyl(ph.ecog, sex) %>% 
  adorn_totals("row") %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns() 
```

#### 2. 

Add variables to this typical [table one](https://thomaselove.github.io/432-notes/building-table-1.html#two-examples-from-the-new-england-journal-of-medicine) created with the [`aresenal`](https://mayoverse.github.io/arsenal/reference/arsenal.html) package

```{r results='asis'}
library(arsenal)

tableby(ph.ecog ~ sex, data = mockdata) %>% 
  summary()
```


#### 3. 

Add a `fig.path` in your \ref{setup} chunk and `knit` again !


#### 4. 

Add a histogram of the ages of the study participants in the plot grid for your report!

First, create, and save a histogram **HINT** check out the [`ggplot2` cheat sheet](https://www.rstudio.com/resources/cheatsheets/)

```{r}

```

Then, add your plot to this grid copied from above. 

```{r}
plot_grid(cow1, p3$plot, 
          label_size = 12,
          labels = c('', "C"),
          ncol = 1)
```

